---
title: "P8130 Final Project"
output: pdf_document
---


### Abstract  
### Introduction (brief context and background of the problem) 
### Methods (data description and statistical methods)  
### Results 
### Conclusions/Discussion  

```{r, message = FALSE}
library(tidyverse)
library(ggplot2)
library(GGally)
library(PerformanceAnalytics)
library(performance)
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Read in dataset
```{r message=FALSE, warning=FALSE}
cdi = read_csv("./cdi.csv") %>% 
  janitor::clean_names()
```

```{r}
## no missing value
cdi %>% 
  select(everything()) %>%  
  summarise_all(funs(sum(is.na(.)))) %>% 
  knitr::kable()
```

## Data cleaning
```{r}
# some normalization for better comparison
cdi = 
  cdi %>% 
  mutate(crm_1000 = crimes/pop*1000,  # as indicated by the project prompt
         docs_1000 = docs/pop*1000,  # every 1000 people how many doctors
         beds_1000 = beds/pop*1000,  # ratio of beds per doctor
         pop_density = pop/area,  # how many people per square miles
         region = factor(region)) %>% 
  dplyr::select(-id, -crimes,-area, -docs, -beds, -totalinc)
```


## Data Exploration
```{r}
## summary statistics, tentative, NOT FINAL
sum_cdi = 
  cdi %>% 
  dplyr::select(crm_1000, docs_1000, pop_density, pop, pop18, pop65, hsgrad, bagrad, poverty,unemp, pcincome,beds_1000)
summary(sum_cdi)
```

```{r}
mean_crm = mean(sum_cdi$crm_1000)
cdi_state = cdi %>% 
  group_by(state) %>% 
  summarize(crime_rate = mean(crm_1000)) %>% 
  mutate(low_high = ifelse(crime_rate>mean_crm, TRUE,FALSE))
  

cdi_state %>% 
  mutate(state = fct_reorder(state, crime_rate)) %>% 
  ggplot(aes(x = state, y = crime_rate))+
  geom_hline(yintercept = mean_crm, color = "red")+
  geom_point(aes(color = low_high),size = 3)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust= 1),
        legend.position = "none")
```

### boxplot for each variable
```{r}
par(mfrow=c(2,3))
boxplot(sum_cdi$crm_1000, main='Crime Rate')
boxplot(sum_cdi$docs_1000, main='Doctor Density')
boxplot(sum_cdi$pop_density,main='Population Density' )
boxplot(sum_cdi$hsgrad, main='High School Graduate')
boxplot(sum_cdi$bagrad, main='Bachelor Graduate')
boxplot(sum_cdi$poverty, main='Poverty')
par(mfrow=c(2,3))
boxplot(sum_cdi$unemp, main='Unemployment Rate')
boxplot(sum_cdi$pcincome, main='Per capita Income')
boxplot(sum_cdi$beds_1000, main='Beds Rate')
boxplot(sum_cdi$pop, main='Population')
boxplot(sum_cdi$pop18, main='PP aged 18-34')
boxplot(sum_cdi$pop65, main='PP aged 65+')
```



### Marginal Correlation and Correlation martix
```{r}
corr_matrix = 
  #cdi %>% 
  #dplyr::select(-state, -region, -cty) %>% 
  sum_cdi %>% 
  chart.Correlation(histogram = TRUE, method = "pearson")
```

### Correlation Heatmap
```{r}
cdi %>% 
  dplyr::select(-state, -region, -cty) %>% 
 # sum_cdi %>% 
  ggcorr(label=TRUE, hjust = 0.9, layout.exp = 2, label_size = 3, label_round = 2)
#corrplot(cor(cdi_1), type = "upper", diag = FALSE)
```

## Build Model

### Backward Elimination
```{r}
mult_fit = lm(crm_1000 ~ ., data = sum_cdi)
summary(mult_fit)
multi_back = step(mult_fit, direction='backward')
```


### Forward Selection
```{r}
multi_forward = step(mult_fit, direction = 'forward')
```

### Both direction
```{r}
multi_both = step(mult_fit, direction = "both")
```

## Residuals vs. Fitted && QQ Plots
```{r}

```


## Check Multicollinearity
```{r}
check_collinearity(multi_forward)
check_collinearity(multi_back)
check_collinearity(multi_both)
```




## Model Diagnostics

### Residuals VS Fitted/Predicted Values

#### To detect unequal error varince and Outliers

```{r}
plot(model1, which = 1)
```

### QQ plot

#### (non) normality of residuals and outliers

```{r}
plot(model1, which = 2)
```

### residuals vs leverage plot

#### To identify influential cases

```{r}

plot(mult.fit1, which = 4)
```

### Box-Cox Transformation

```{r}
boxcox(model1, lambda = seq(-3, 3, by = 0.25))
```



## Vairable Selection

