---
title: "P8130 Final Project"
geometry: margin=1.8cm
output: pdf_document
---


### Abstract  
### Introduction (brief context and background of the problem) 
### Methods (data description and statistical methods)  
### Results 
### Conclusions/Discussion  

```{r, message = FALSE}
# import necessary datasets
library(tidyverse)
library(ggplot2)
library(GGally)
library(PerformanceAnalytics)
library(performance)
library(MASS)
library(leaps)
library(modelr)
library(olsrr)
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Read in dataset
```{r message=FALSE, warning=FALSE}
cdi = read_csv("./cdi.csv") %>% 
  janitor::clean_names()
```

```{r}
## no missing value
cdi %>% 
  dplyr::select(everything()) %>%  
  summarise_all(funs(sum(is.na(.)))) %>% 
  knitr::kable()
```

## Data cleaning
First, some normalization for better comparison
```{r}
cdi = 
  cdi %>% 
  mutate(crm_1000 = crimes/pop*1000,  # as indicated by the project prompt
         docs_1000 = docs/pop*1000,  # every 1000 people how many doctors
         beds_1000 = beds/pop*1000, # similar as above
         pop_density = pop/area,  # how many people per square miles
         northeast = ifelse(region == 1, 1, 0),  # region as dummy varaible
         northcentral = ifelse(region == 2, 1, 0),
         south = ifelse(region == 3, 1, 0)) %>% 
  dplyr::select(-id, -crimes,-area, -docs, -beds, -totalinc, -region)
```


## Data Exploration
```{r}
## summary statistics, tentative, NOT FINAL
sum_cdi = 
  cdi %>% 
  dplyr::select(-c(cty, state))
summary(sum_cdi)
```

```{r}
mean_crm = mean(sum_cdi$crm_1000)
cdi_state = cdi %>% 
  group_by(state) %>% 
  summarize(crime_rate = mean(crm_1000)) %>% 
  mutate(low_high = ifelse(crime_rate>mean_crm, TRUE,FALSE))
  

cdi_state %>% 
  mutate(state = fct_reorder(state, crime_rate)) %>% 
  ggplot(aes(x = state, y = crime_rate))+
  geom_hline(yintercept = mean_crm, color = "red")+
  geom_point(aes(color = low_high),size = 3)+
  ggtitle("Crime Rates in the US States") +
  labs(y = "Crime Rate Per 1000", x = "States") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust= 1),
        legend.position = "none")
```

### boxplot for each variable
```{r}
par(mfrow=c(2,3))
boxplot(sum_cdi$crm_1000, main='Crime Rate')
boxplot(sum_cdi$docs_1000, main='Doctor Density')
boxplot(sum_cdi$pop_density,main='Population Density' )
boxplot(sum_cdi$hsgrad, main='High School Graduate')
boxplot(sum_cdi$bagrad, main='Bachelor Graduate')
boxplot(sum_cdi$poverty, main='Poverty')
par(mfrow=c(2,3))
boxplot(sum_cdi$unemp, main='Unemployment Rate')
boxplot(sum_cdi$pcincome, main='Per capita Income')
boxplot(sum_cdi$beds_1000, main='Beds Rate')
boxplot(sum_cdi$pop, main='Population')
boxplot(sum_cdi$pop18, main='PP aged 18-34')
boxplot(sum_cdi$pop65, main='PP aged 65+')
```

### Histogram for each variable
```{r}
par(mfrow=c(2,3))
hist(sum_cdi$crm_1000, main='Crime Rate')
hist(sum_cdi$docs_1000, main='Doctor Density')
hist(sum_cdi$pop_density,main='Population Density' )
hist(sum_cdi$hsgrad, main='High School Graduate')
hist(sum_cdi$bagrad, main='Bachelor Graduate')
hist(sum_cdi$poverty, main='Poverty')
par(mfrow=c(2,3))
hist(sum_cdi$unemp, main='Unemployment Rate')
hist(sum_cdi$pcincome, main='Per capita Income')
hist(sum_cdi$beds_1000, main='Beds Rate')
hist(sum_cdi$pop, main='Population')
hist(sum_cdi$pop18, main='PP aged 18-34')
hist(sum_cdi$pop65, main='PP aged 65+')
```

### Marginal Correlation and Correlation martix
```{r}
corr_matrix = 
  cdi %>% 
  dplyr::select(-state, -cty, -northeast, -northcentral, -south) %>% 
  chart.Correlation(histogram = TRUE, method = "pearson")
```

### Correlation Heatmap
```{r}
cdi %>% 
  dplyr::select(-state, -cty, -northeast, -northcentral, -south) %>% 
  dplyr::select(crm_1000, everything()) %>% 
  ggcorr(label=TRUE, hjust = 0.9, layout.exp = 2, label_size = 3, label_round = 2)
```

## Build Model

### Full Model
Let's start with the full model
```{r}
mult_fit = lm(crm_1000 ~ ., data = sum_cdi)
summary(mult_fit)
```

Model diagnostics of the full model
```{r}
par(mfrow=c(2,2))
plot(mult_fit)
```

```{r}
# get the lambda for the transformation
bc_model = boxcox(mult_fit, lambda = seq(-3, 3, by = 0.25))
lamb = bc_model$x[which.max(bc_model$y)]
lamb
```

~0.5, thus we applied square root to the Y. Also we get rid of the influential points. The full model is the basis of other models, thus we choose to filter the outliers out at first.
```{r}
sum_cdi_mod = sum_cdi[-c(1,6),]  # filter out outlier and store it as the new dataset
full_trans_fit = lm(sqrt(crm_1000) ~.,data = sum_cdi_mod)  # refit
```

```{r}
# check again
par(mfrow=c(2,2))
plot(full_trans_fit)
summary(full_trans_fit)
check_collinearity(full_trans_fit)
```

We will just use the transformed models for the further model fits

### Backward Elimination
```{r}
multi_back = step(full_trans_fit, direction='backward')
multi_back
```
sqrt(crm_1000) ~ pop + pop18 + hsgrad + bagrad + poverty + unemp+ pcincome + beds_1000 + pop_density + northeast + northcentral + south, data = sum_cdi_mod


Model Diagnostic
```{r}
par(mfrow = c(2,2))
plot(multi_back)
check_collinearity(multi_back)
```


### Forward Selection
```{r}
multi_forward = step(full_trans_fit, direction = 'forward')
multi_forward
```
sqrt(crm_1000) ~ pop + pop18 + pop65 + hsgrad + bagrad + poverty + unemp + pcincome + docs_1000 + beds_1000 + pop_density + northeast + northcentral + south, data = sum_cdi_mod

Forward selection generated the same result as the full model, thus we will not consider it from now on. 


### Both direction
```{r}
multi_both = step(full_trans_fit, direction = "both")
multi_both
```

sqrt(crm_1000) ~ pop + pop18 + hsgrad + bagrad + poverty + unemp + pcincome + beds_1000 + pop_density + northeast + northcentral + south, data = sum_cdi_mod

the same model as the backward selection, we will not focus on this model from now on
    

### Interaction Model

Choose backward selection-generated model as the basis since it's with the highest adjusted r-squared. The choice of interaction is somewhat arbitrary, mainly based on the correlation heatmap and also to avoid high collinearity. After a few tries, we added two more interaction terms as the following
```{r}
multi_interact = lm(sqrt(crm_1000) ~ pop + pop18 + hsgrad + bagrad + poverty + unemp+ pcincome + beds_1000 
                    + pop_density + northeast + northcentral + south + pop*bagrad, data = sum_cdi_mod)
summary(multi_interact)
```

```{r}
anova(multi_back, multi_interact)
```

Model diagnostic
```{r}
par(mfrow = c(2,2))
plot(multi_interact)
check_collinearity(multi_interact)
```

## Test based procedures

### Model diagnostics
```{r}
b = regsubsets(sqrt(crm_1000)~ ., data = sum_cdi_mod, nvmax = 15)
rs = summary(b)

# plot of Cp and Adj-R2 as functions of parameters
par(mfrow=c(1,2))

plot(2:15, rs$cp, xlab="No of parameters", ylab="Cp Statistic")
abline(0,1)

plot(2:15, rs$adjr2, xlab="No of parameters", ylab="Adj R2")
```


Adjusted R-squared based model
```{r}
models_generator = function(predict_num, models){
  predict_intent = summary(models)$which[predict_num, -1]
  predict = names(which(predict_intent == TRUE))
  predictors = paste(predict, collapse = " + ")
  text = paste0("sqrt(crm_1000) ~ ", predictors)
  return(text)
}
adjr2_num = which.max(rs$adjr2)
models_generator(adjr2_num, b)
```


Cp based model 
```{r}
cp_num = which.min(rs$cp)
models_generator(cp_num, b)
```

Fit both models
```{r}
# r-adj square
multi_r_adj = lm(sqrt(crm_1000) ~ pop + pop18 + hsgrad + bagrad + poverty + unemp + pcincome + beds_1000 + pop_density + northeast + northcentral + south, data = sum_cdi_mod)
summary(multi_r_adj)
```

Model diagnostic
```{r}
par(mfrow = c(2,2))
plot(multi_r_adj)
check_collinearity(multi_r_adj)
```

```{r}
# cp value based
multi_cp = lm(sqrt(crm_1000) ~ pop + pop18 + bagrad + poverty + pcincome + beds_1000 + pop_density + northeast + northcentral, data = sum_cdi_mod)
summary(multi_cp)
```

Model diagnostic
```{r}
par(mfrow = c(2,2))
plot(multi_cp)
check_collinearity(multi_cp)
```


```{r}
a_row = function(model_data){
  model_data %>% 
  broom::glance() %>% 
  dplyr::select(adj.r.squared, AIC, BIC)
}
add_in = rbind(
  ols_mallows_cp(full_trans_fit,full_trans_fit),
  ols_mallows_cp(multi_back,full_trans_fit),
  ols_mallows_cp(multi_interact,full_trans_fit),
  ols_mallows_cp(multi_r_adj,full_trans_fit),
  ols_mallows_cp(multi_cp,full_trans_fit))
rmse_add = rbind(
  rmse(full_trans_fit, data = sum_cdi_mod),
  rmse(multi_back, data = sum_cdi_mod),
  rmse(multi_interact, data = sum_cdi_mod),
  rmse(multi_r_adj, data = sum_cdi_mod),
  rmse(multi_cp, data = sum_cdi_mod)
)
rbind(a_row(full_trans_fit),
      a_row(multi_back),
      a_row(multi_interact),
      a_row(multi_r_adj),
      a_row(multi_cp))%>% 
  mutate(model = c("Full model", "Backward Selection", "Interaction", "Adj R Based", "Cp Value Based"),
         cp = add_in,
         rmse = rmse_add) %>% 
  relocate(model)

```

## Cross Validation
```{r}
set.seed(1)
  
cv_df = 
  crossv_kfold(sum_cdi_mod, k=10) %>%   # k-fold = 5
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )
cv_df = 
  cv_df %>% 
  mutate(
    full_fit = map(.x = train, ~lm(sqrt(crm_1000) ~., data = .x)),
    back_fit = map(.x = train, ~lm(sqrt(crm_1000) ~ pop + pop18 + hsgrad + bagrad + poverty + 
                                     unemp+ pcincome + beds_1000 + pop_density + northeast + 
                                     northcentral + south,data = .x)),
    interact_fit = map(.x = train, ~lm(sqrt(crm_1000) ~ pop + pop18 + hsgrad + bagrad + poverty + unemp+ pcincome 
                                       + beds_1000 + pop_density + northeast + 
                                         northcentral + south + pop*bagrad, data = .x)),
    adj_fit = map(.x = train, ~lm(sqrt(crm_1000) ~ pop + pop18 + hsgrad + bagrad + poverty + unemp +
                                    pcincome + beds_1000 + pop_density + northeast + northcentral + south, data = .x)),
    cp_fit = map(.x = train, ~lm(sqrt(crm_1000) ~ pop + pop18 + bagrad + poverty + pcincome +
                                   beds_1000 + pop_density + northeast + northcentral, data = .x))
  ) %>% 
  mutate(
    rmse_full = map2_dbl(.x = full_fit, .y = test, ~rmse(model = .x, data = .y)),
    rmse_back = map2_dbl(.x = back_fit, .y = test, ~rmse(model = .x, data = .y)),
    rmse_interact = map2_dbl(.x = interact_fit, .y = test, ~rmse(model = .x, data = .y)),
    rmse_adj = map2_dbl(.x = adj_fit, .y = test, ~rmse(model = .x, data = .y)),
    rmse_cp = map2_dbl(.x = cp_fit, .y = test, ~rmse(model = .x, data = .y)),
  )
```

```{r}
cv_df %>% 
  dplyr::select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) %>% 
  mutate(model = fct_relevel(model, "full", "back", "interact", "Adj R-squared", "Cp")) %>% 
  ggplot(aes(x = model, y = rmse,))+
  geom_violin(aes(fill = model), alpha = 0.3)+
  scale_x_discrete(labels = c("Full model", "Backward Selection", "Interaction", "Adj R-Squared Based", "Cp Value Based"))+
  ggtitle("RMSE Distribution Plots") +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "none")  +  # the display of legends is redundant
  labs(y = "RMSE", x = "Models")
```

