---
title: "P8130 Final Project"
geometry: margin=1.8cm
output: pdf_document
---


### Abstract  
### Introduction (brief context and background of the problem) 
### Methods (data description and statistical methods)  
### Results 
### Conclusions/Discussion  

```{r, message = FALSE}
# import necessary datasets
library(tidyverse)
library(ggplot2)
library(GGally)
library(PerformanceAnalytics)
library(performance)
library(MASS)
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Read in dataset
```{r message=FALSE, warning=FALSE}
cdi = read_csv("./cdi.csv") %>% 
  janitor::clean_names()
```

```{r}
## no missing value
cdi %>% 
  dplyr::select(everything()) %>%  
  summarise_all(funs(sum(is.na(.)))) %>% 
  knitr::kable()
```

## Data cleaning
First, some normalization for better comparison
```{r}
cdi = 
  cdi %>% 
  mutate(crm_1000 = crimes/pop*1000,  # as indicated by the project prompt
         docs_1000 = docs/pop*1000,  # every 1000 people how many doctors
         beds_1000 = beds/pop*1000, # similar as above
         pop_density = pop/area,  # how many people per square miles
         northeast = ifelse(region == 1, 1, 0),  # region as dummy varaible
         northcentral = ifelse(region == 2, 1, 0),
         south = ifelse(region == 3, 1, 0)) %>% 
  dplyr::select(-id, -crimes,-area, -docs, -beds, -totalinc, -region)
```


## Data Exploration
```{r}
## summary statistics, tentative, NOT FINAL
sum_cdi = 
  cdi %>% 
  dplyr::select(-c(cty, state))
summary(sum_cdi)
```

```{r}
mean_crm = mean(sum_cdi$crm_1000)
cdi_state = cdi %>% 
  group_by(state) %>% 
  summarize(crime_rate = mean(crm_1000)) %>% 
  mutate(low_high = ifelse(crime_rate>mean_crm, TRUE,FALSE))
  

cdi_state %>% 
  mutate(state = fct_reorder(state, crime_rate)) %>% 
  ggplot(aes(x = state, y = crime_rate))+
  geom_hline(yintercept = mean_crm, color = "red")+
  geom_point(aes(color = low_high),size = 3)+
  ggtitle("Crime Rates in the US States") +
  labs(y = "Crime Rate Per 1000", x = "States") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust= 1),
        legend.position = "none")
```

### boxplot for each variable
```{r}
par(mfrow=c(2,3))
boxplot(sum_cdi$crm_1000, main='Crime Rate')
boxplot(sum_cdi$docs_1000, main='Doctor Density')
boxplot(sum_cdi$pop_density,main='Population Density' )
boxplot(sum_cdi$hsgrad, main='High School Graduate')
boxplot(sum_cdi$bagrad, main='Bachelor Graduate')
boxplot(sum_cdi$poverty, main='Poverty')
par(mfrow=c(2,3))
boxplot(sum_cdi$unemp, main='Unemployment Rate')
boxplot(sum_cdi$pcincome, main='Per capita Income')
boxplot(sum_cdi$beds_1000, main='Beds Rate')
boxplot(sum_cdi$pop, main='Population')
boxplot(sum_cdi$pop18, main='PP aged 18-34')
boxplot(sum_cdi$pop65, main='PP aged 65+')
```



### Marginal Correlation and Correlation martix
```{r}
corr_matrix = 
  cdi %>% 
  dplyr::select(-state, -cty, -northeast, -northcentral, -south) %>% 
  chart.Correlation(histogram = TRUE, method = "pearson")
```

### Correlation Heatmap
```{r}
cdi %>% 
  dplyr::select(-state, -cty, -northeast, -northcentral, -south) %>% 
  dplyr::select(crm_1000, everything()) %>% 
  ggcorr(label=TRUE, hjust = 0.9, layout.exp = 2, label_size = 3, label_round = 2)
```
## Build Model
### Full Model
Let's start with the full model
```{r}
mult_fit = lm(crm_1000 ~ ., data = sum_cdi)
summary(mult_fit)
```

Model diagnostics of the full model
```{r}
par(mfrow=c(2,2))
plot(mult_fit)
```

```{r}
# get the lambda for the transformation
bc_model = boxcox(mult_fit, lambda = seq(-3, 3, by = 0.25))
lamb = bc_model$x[which.max(bc_model$y)]
lamb
```

~0.5, thus we applied square root to the Y. Also we get rid of the influential points. The full model is the basis of other models, thus we choose to filter the outliers out at first.
```{r}
sum_cdi_mod = sum_cdi[-c(1,6),]  # filter out outlier and store it as the new dataset
full_trans_fit = lm(sqrt(crm_1000) ~.,data = sum_cdi_mod)  # refit
```

```{r}
# check again
par(mfrow=c(2,2))
plot(full_trans_fit)
summary(full_trans_fit)
check_collinearity(full_trans_fit)
```

We will just use the transformed models for the further model fits

### Backward Elimination
```{r}
multi_back = step(full_trans_fit, direction='backward')
multi_back
```
sqrt(crm_1000) ~ pop + pop18 + hsgrad + bagrad + poverty + unemp+ pcincome + beds_1000 + pop_density + northeast + northcentral + south, data = sum_cdi


Model Diagnostic
```{r}
par(mfrow = c(2,1))
plot(multi_back, which = 1)
plot(multi_back, which = 4)
```

```{r}
# get rid of the outliers first
back_without = sum_cdi[-c(1),]

with_back = lm(sqrt(crm_1000) ~ pop + pop18 + hsgrad + bagrad + poverty + unemp+
    pcincome + beds_1000 + pop_density + northeast + northcentral + 
    south, data = sum_cdi) 
without_back = lm(sqrt(crm_1000) ~ pop + pop18 + hsgrad + bagrad + poverty + unemp+
    pcincome + beds_1000 + pop_density + northeast + northcentral + 
    south, data = back_without)
summary(with_back); summary(without_back)
check_collinearity(without_back)
```


### Forward Selection
```{r}
multi_forward = step(full_trans_fit, direction = 'forward')
multi_forward
```
sqrt(crm_1000) ~ pop + pop18 + pop65 + hsgrad + bagrad + poverty + unemp + pcincome + docs_1000 + beds_1000 + pop_density + northeast + northcentral + south, data = sum_cdi_mod

Model diagnostic
```{r}
par(mfrow = c(2,1))
plot(multi_forward, which = 1)
plot(multi_forward, which = 4)
```

```{r}
forward_without = sum_cdi[-c(1,416),]
with_for = lm(sqrt(crm_1000) ~ pop + pop18 + pop65 + hsgrad + 
    bagrad + poverty + unemp + pcincome + docs_1000 + beds_1000 + 
    pop_density + northeast + northcentral + south, data = sum_cdi_mod) 
without_for = lm(sqrt(crm_1000) ~ pop + pop18 + pop65 + hsgrad + 
    bagrad + poverty + unemp + pcincome + docs_1000 + beds_1000 + 
    pop_density + northeast + northcentral + south,data = forward_without)
summary(with_for); summary(without_for)
check_collinearity(without_for)
```


### Both direction
```{r}
multi_both = step(full_trans_fit, direction = "both")
multi_both
```

sqrt(crm_1000) ~ pop + pop18 + hsgrad + bagrad + poverty + unemp + pcincome + beds_1000 + pop_density + northeast + northcentral + south, data = sum_cdi_mod
    
Model diagnostic
```{r}
par(mfrow = c(2,1))
plot(multi_both, which = 1)
plot(multi_both, which = 4)
```

```{r}
both_without = sum_cdi[-c(1),]

with_both = lm(sqrt(crm_1000) ~ pop + pop18 + hsgrad + bagrad + 
    poverty + unemp + pcincome + beds_1000 + pop_density + northeast + 
    northcentral + south, data = sum_cdi_mod) 
without_both = lm(sqrt(crm_1000) ~ pop + pop18 + hsgrad + bagrad + 
    poverty + unemp + pcincome + beds_1000 + pop_density + northeast + 
    northcentral + south, data = both_without)
summary(with_both); summary(without_both)
check_collinearity(without_both)
```
